---
title: "Introduction to Bayesian regression models"
subtitle: "02 - Bayesian regression"
author: "Stefano Coretta"
institute: "University of Edinburgh"
editor: source
format:
  mono-light-revealjs:
    theme: [default, custom.scss]
    history: false
    preview-links: true
filters:
  - tachyonsextra
execute: 
  echo: true
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
theme_set(theme_minimal())

library(lme4)
library(lmerTest)
library(ggeffects)
library(brms)
library(tidybayes)
library(bayesplot)
```

## Regression models

::: box-note
-   Simplest regression model:

$$
y \sim \beta_0 + \beta_1 \cdot x
$$

-   where $y$ and $x$ are numeric variables.

-   <https://stefanocoretta.shinyapps.io/lines/>
:::

## MALD

::: box-tip
**Does the mean phone-level distance affect reaction times**?
:::

. . .

::: box-note
-   Mean phone-level distance is a proxy of word "uniqueness".

-   More unique words should take longer to recognise as real words because there are less competitors (i.e. similar words) that can have a facilitatory effect on recognition.
:::

## MALD: `RT` and `PhonLev`

```{r}
#| label: mald
#| echo: false

mald <- readRDS("./data/mald.rds")
```

```{r}
#| label: fig-dist-stat
#| fig-asp: 0.75
#| echo: false
#| fig-cap: "Scatter plot of mean phone-level distance vs reaction times."

mald |> 
  ggplot(aes(PhonLev, RT)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", formula = y ~ x) +
  labs(
    x = "Phone-level distance", y = "RT (ms)"
  )
```

## MALD: regression

$$
RT \sim \beta_0 + \beta_1 \cdot \text{PhonLev}
$$

. . .

or

$$
\begin{align}
RT & \sim \beta_0 + \beta_1 \cdot \text{PhonLev} + \epsilon\\
\epsilon & \sim Gaussian(0, \sigma)
\end{align}
$$

. . .

or

$$
\begin{align}
RT & \sim Gaussian(\mu, \sigma)\\
\mu & = \beta_0 + \beta_1 \cdot \text{PhonLev}\\
\end{align}
$$

::: box-note
The model estimates the posterior probability distributions of $\beta_0, \beta_1, \sigma$.
:::

## What are your expectations?

```{=html}
<iframe allowfullscreen frameborder="0" height="100%" mozallowfullscreen style="min-width: 500px; min-height: 355px" src="https://app.wooclap.com/events/JEJPPA/questions/68346b14678ad4313b2e2b05" width="100%"></iframe>
```

## MALD: regression

```{r}
#| label: brm-lev

brm_lev <- brm(
  RT ~ 1 + PhonLev,
  data = mald,
  family = "gaussian",
  file = "./data/cache/brm_lev.rds",
  chains = 4, 
  iter = 2000,
  cores = 4
)

summary(brm_lev)
```

## Interpret the summary: `Intercept`

```{r}
#| label: brm-lev-summ-int
#| echo: false

cat(capture.output(summary(brm_lev))[8:11], sep = "\n")
```

::: box-warning
`Intercept` is $\beta_0$.

-   $\beta_0$ is the expected mean RT when `PhonLev` is 0.

-   The posterior probability distribution of $\beta_0$ is $\mathcal{P}(862, 35)$.

-   There is a **95% probability** that the mean RTs when `PhonLev` is 0 is **between 793 and 929 ms**, conditional on the model and data. This is the 95% Credible Interval (CrI) of $\mathcal{P}(862, 35)$.
:::

## Interpret the summary: `PhonLev`

```{r}
#| label: brm-lev-summ-phonlev
#| echo: false

cat(capture.output(summary(brm_lev))[8:11], sep = "\n")
```

::: box-warning
`PhonLev` is $\beta_1$.

-   $\beta_1$ is the expected difference in RT *for each unit change* of `PhonLev`.

-   The posterior probability distribution of $\beta_1$ is $\mathcal{P}(26, 5)$.

-   There is a **95% probability** that each unit change of `PhonLev` corresponds to a change in mean RT **between +17 and +34 ms**, conditional on the model and data. This is the 95% Credible Interval (CrI) of $\mathcal{P}(26, 5)$.
:::

## Effect of `PhonLev` on RTs

```{r}
#| label: fig-brm-lev
#| fig-asp: 0.75
#| fig-cap: "Predicted effect of mean phone-level distance on RTs from a Bayesian regression model."

conditional_effects(brm_lev, "PhonLev")
```

## MCMC draws

```{r}
#| label: brm-lev-draws

brm_lev_draws <- as_draws_df(brm_lev)

brm_lev_draws
```

## Predicted RTs at representative values of `PhonLev`

```{r}
#| label: brm-lev-draws-2

range(mald$PhonLev)

brm_lev_draws <- brm_lev_draws |> 
  mutate(
    # RTs when PhonLev is 5
    rt_05 = b_Intercept + b_PhonLev * 5,
    # RTs when PhonLev is 10
    rt_10 = b_Intercept + b_PhonLev * 10,
    # RTs when PhonLev is 15
    rt_15 = b_Intercept + b_PhonLev * 15,
  )
```

## Predicted RTs at representative values of `PhonLev`: plot

```{r}
#| label: fig-lev-5-10-15
#| output-location: slide
#| fig-cap: "Predicted RTs at representative values of mean phon-level distance."

library(ggdist)

brm_lev_draws |> 
  select(rt_05, rt_10, rt_15) |> 
  
  # Pivot so to make plotting easier
  pivot_longer(everything()) |> 
  # Pivot creates a name column with values `rt_05, rt_10, rt_15`
  # and a value column with the draws values
  
  ggplot(aes(value, name)) +
  # Stat halfeye from ggdist
  stat_halfeye() +
  labs(
    caption = "Point = median, thick line = 66% CrI, thin line = 95% CrI."
  )
```

## Calculating CrIs from the draws

```{r}
#| label: brm-lev-cri
#| warning: false

library(posterior)

brm_lev_cri <- brm_lev_draws |> 
  select(rt_05, rt_10, rt_15) |> 
  pivot_longer(everything()) |> 
  group_by(name) |> 
  summarise(
    # Use quantile2() from posterior package
    lo_95 = round(quantile2(value, 0.025)),
    hi_95 = round(quantile2(value, 0.0975))
  )

brm_lev_cri
```

## Width of the CrIs

```{r}
#| label: brm-lev-cri-width

brm_lev_cri |> 
  mutate(
    cri_width = hi_95 - lo_95
  )
```

## Reporting

> We fitted a Bayesian Gaussian regression model with brms [@burkner2017] in R 4.5.0 [@rcoreteam2025]. The model had reaction times as the outcome variable and mean phone-level distance as the only predictor. The model was fitted with 4 MCMC chains and 2000 iterations per chains. The default brms priors were used.
>
> According to the model results, for every increase of mean phone-level distance by 1 unit, reaction times increase by 17-34 ms, at 95% probability (\\(\\beta\\) = 26.05, SD = 4.85). The 95% CrIs of the predicted reaction times at representative values of mean distance are: 968-976 ms for mean distance = 5, 1092-1103 ms for mean distance = 10, 1177-1203 for mean distance = 15.

## References
