---
title: "Introduction to Bayesian regression models"
subtitle: "04 - Categorical predictors"
author: "Stefano Coretta"
institute: "University of Edinburgh"
editor: source
format:
  mono-light-revealjs:
    theme: [default, custom.scss]
    history: false
filters:
  - tachyonsextra
execute: 
  echo: true
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
theme_set(theme_minimal())

library(lme4)
library(ggeffects)
library(brms)
library(tidybayes)
library(bayesplot)
```

## MALD again

```{r}
#| label: fig-mald-isword
#| fig-asp: 0.75
#| fig-cap: "RTs by lexical status."
#| fig-align: "center"
#| echo: false

mald <- readRDS("./data/mald.rds")

mald |> 
  filter(RT > 500) |> 
  ggplot(aes(IsWord, RT, fill = IsWord)) +
  geom_jitter(width = 0.2, alpha = 0.2) +
  geom_violin(width = 0.1) +
  theme(legend.position = "none")
```

## MALD: `IsWord`

::: box-tip
**Does the lexical status of the target word (real or nonce) affect reaction times**?
:::

## Categorical predictors: treatment contrasts

```{r}
#| label: brm-cat

brm_cat <- brm(
  RT ~ 1 + PhonLev + IsWord,
  data = mald,
  family = lognormal,
  cores = 4,
  seed = 9812,
  file = "data/cache/brm_cat.rds"
)

brm_cat
```

## Interpret treatment contrasts

```{r}
#| label: brm-cat-summ
#| echo: false

cat(capture.output(summary(brm_cat))[8:12], sep = "\n")
```

::: box-note
-   `Intercept`: mean RT when `PhonLev` is 0 and `IsWord` = TRUE.

-   `PhonLev`: change in mean RT for each unit increase in `PhonLev` (average across `IsWord`).

-   `IsWordFalse`: difference in mean RT between `IsWord` = FALSE and TRUE (when `PhonLev` = 0).
:::

. . .

::: box-warning
So, does the lexical status of the word affect RTs?
:::

## Indexing categorical predictors

$$
\begin{align}
RT_i & \sim Lognormal(\mu_i, \sigma)\\
log(\mu_i) & = \beta_1 \cdot W_{\text{T}[i]} + \beta_2 \cdot W_{\text{F}[i]} + \beta_3 \cdot PL_i\\
\end{align}
$$

. . .

::: box-note
-   $\beta_1$: mean RT when `IsWord` = TRUE ($W_{T}$).

-   $\beta_1$: mean RT when `IsWord` = FALSE ($W_{F}$).

-   $\beta_3$: change in RT for each unit increase of `PhonLev` ($PL$).
:::

## Indexing with intercept suppression

```{r}
#| label: brm-ind

brm_ind <- brm(
  RT ~ 0 + IsWord + PhonLev,
  data = mald,
  family = lognormal,
  cores = 4,
  seed = 6293,
  file = "data/cache/brm_ind.rds"
)

brm_ind
```

## RTs by lexical status (at mean `PhonLev`)

```{r}
#| label: brm-ind-cond-mean

conditional_effects(brm_ind, "IsWord")
```

## RTs by lexical status (`PhonLev` = 0)

```{r}
#| label: brm-ind-cond-0
#| fig-asp: 0.75

conditional_effects(brm_ind, "IsWord", conditions = list(PhonLev = 0))
```

## RTs by lexical status (`PhonLev` = 5)

```{r}
#| label: brm-ind-cond-5

conditional_effects(brm_ind, "IsWord", conditions = list(PhonLev = 5))
```

## RTs by lexical status (`PhonLev` = 10)

```{r}
#| label: brm-ind-cond-10

conditional_effects(brm_ind, "IsWord", conditions = list(PhonLev = 10))
```

## RTs by lexical status (`PhonLev` = 14)

```{r}
#| label: brm-ind-cond-14

conditional_effects(brm_ind, "IsWord", conditions = list(PhonLev = 14))
```

## Calculate RT difference from the MCMC draws: calculate RTs

```{r}
#| label: brm-ind-draws

brm_ind_draws <- as_draws_df(brm_ind) |> 
  mutate(
    # RT when PhonLev = 5 (min PhonLev)
    rt_true_5 = exp(b_IsWordTRUE + b_PhonLev * 5),
    rt_false_5 = exp(b_IsWordFALSE + b_PhonLev * 5),
    # RT when PhonLev = 7 (mean PhonLev)
    rt_true_7 = exp(b_IsWordTRUE + b_PhonLev * 7),
    rt_false_7 = exp(b_IsWordFALSE + b_PhonLev * 7),
    # RT when PhonLev = 14 (max PhonLev)
    rt_true_14 = exp(b_IsWordTRUE + b_PhonLev * 14),
    rt_false_14 = exp(b_IsWordFALSE + b_PhonLev * 14),
  )
```

## Calculate RT difference from the MCMC draws: get difference

```{r}
#| label: brm-ind-draws-diff

brm_ind_draws <- brm_ind_draws |> 
  mutate(
    rt_05_diff = rt_false_5 - rt_true_5,
    rt_07_diff = rt_false_7 - rt_true_7,
    rt_14_diff = rt_false_14 - rt_true_14,
  )

brm_ind_draws |> select(rt_05_diff:rt_14_diff)
```

## Calculate CrIs of difference

```{r}
#| label: brm-ind-cris

library(posterior)

brm_ind_draws |> 
  select(rt_05_diff:rt_14_diff) |> 
  pivot_longer(everything()) |> 
  group_by(name) |> 
  summarise(
    # Use quantile2() from posterior package
    lo_95 = round(quantile2(value, 0.025)),
    hi_95 = round(quantile2(value, 0.0975))
  )
```

## MALD: `IsWord`

::: box-tip
**Does the lexical status of the target word (real or nonce) affect reaction times**?
:::

. . .

::: box-note
Yes.
:::

## But...

```{r}
#| label: fig-mald-pl-w
#| fig-asp: 0.75
#| fig-cap: "RTs by lexical status and mean phone-level distance."
#| fig-align: "center"
#| echo: false

mald |> 
  filter(RT > 500) |> 
  ggplot(aes(PhonLev, RT_log)) +
  geom_point(alpha = 0.05) +
  geom_smooth(aes(colour = IsWord), method = "lm", formula = "y~x")
```

