---
title: "Introduction to Bayesian Linear Models"
subtitle: "04 - Interactions"
author: "Stefano Coretta"
institute: "University of Edinburgh"
editor: source
format:
  mono-light-revealjs:
    theme: [default, custom.scss]
    history: false
filters:
  - tachyonsextra
execute: 
  echo: true
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
theme_set(theme_minimal())

library(ggeffects)
library(brms)
library(tidybayes)
library(bayesplot)
library(HDInterval)
library(truncdist)
library(marginaleffects)
my_seed <- 0745
```

```{r}
#| label: mald
#| echo: false

mald <- readRDS("./data/mald.rds")
```

## Interactions: the traditional way

```{r}
#| label: brm-4

brm_4 <- brm(
  # RT ~ IsWord + PhoneLev + IsWord:PhonLev
  RT ~ IsWord * PhonLev,
  family = gaussian,
  data = mald,
  cores = 4,
  file = "data/cache/brm_4",
  seed = my_seed
)
```


## Interactions: the traditional way (model summary)

```{r}
#| label: brm-4-summ

brm_4
```

## Interactions: no intercept

```{r}
#| label: brm-4a

brm_4a <- brm(
  RT ~ 0 + IsWord + IsWord:PhonLev,
  family = gaussian,
  data = mald,
  cores = 4,
  file = "data/cache/brm_4a",
  seed = my_seed
)
```


## Interactions: no intercept (model summary)

```{r}
#| label: brm-4a-summ

brm_4a
```

## Conditional posterior probabilities

```{r}
#| label: brm-4a-cond

conditional_effects(brm_4a, "PhonLev:IsWord")
```

## Conditional posterior probabilities: spaghetti

```{r}
#| label: brm-4a-cond-2

brm_4a_condp <- conditional_effects(brm_4a, "PhonLev:IsWord", spaghetti = TRUE, ndraws = 100)
plot(brm_4a_condp, plot = FALSE, spaghetti_args = list(linewidth = 2))[[1]] +
  theme(
    plot.background = element_rect(fill = "black"),
    text = element_text(colour = "white")
  )
```

## Conditional posteriors at specific values

```{r}
#| label: brm-4a-cond-3
conds <- tibble(
  PhonLev = c(7.5, 10),
  cond__ = c("PhonLev = 7.5", "PhonLev = 10")
)

conditional_effects(brm_4a, "IsWord", conditions = conds)
```

## Comparisons (aka contrasts) at specific values

```{r}
#| label: brm-4a-comp

# library(marginaleffects)

comparisons(brm_4a, variables = "IsWord", newdata = datagrid(PhonLev = c(7.5, 10, 12)))
```

## Comparisons (aka contrasts): 70% CrI

```{r}
#| label: brm-4a-comp-70

comparisons(brm_4a, variables = "IsWord", newdata = datagrid(PhonLev = c(7.5, 10, 12)), conf_level = 0.7)
```

## Comparisons (aka contrasts) along `PhonLev`

```{r}
#| label: brm-4a-comp-phonlev

plot_comparisons(brm_4a, variables = "IsWord", condition = "PhonLev", conf_level = 0.7)
```


## Posterior predictive checks

```{r}
#| label: brm-4a-pp

pp_check(brm_4a, ndraws = 100)
```

