---
title: "Introduction to Bayesian regression models"
subtitle: "05 - Interactions"
author: "Stefano Coretta"
institute: "University of Edinburgh"
editor: source
format:
  mono-light-revealjs:
    theme: [default, custom.scss]
    history: false
filters:
  - tachyonsextra
execute: 
  echo: true
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
theme_set(theme_minimal())

library(lme4)
library(ggeffects)
library(brms)
library(tidybayes)
library(bayesplot)
```

## MALD

```{r}
#| label: fig-mald-pl-w
#| fig-asp: 0.75
#| fig-cap: "RTs by lexical status and mean phone-level distance."
#| fig-align: "center"
#| echo: false

mald <- readRDS("./data/mald.rds")

mald |> 
  filter(RT > 500) |> 
  ggplot(aes(PhonLev, RT_log)) +
  geom_point(alpha = 0.05) +
  geom_smooth(aes(colour = IsWord), method = "lm", formula = "y~x")
```

## MALD: `IsWord` and `PhonLev`

::: box-tip
**Research questions**

1.  Does the lexical status of the word affect RTs?

2.  Does the mean phone-level distance affect RTs?

3.  Does the effect of mean phone-level distance differ depending on the lexical status?
:::

. . .

::: box-warning
For pedagogical reasons we fitted many regression models so far, but you should fit a single model to answer these questions.
:::

## Interactions

$$
\begin{align}
RT_i & \sim Lognormal(\mu_i, \sigma_i)\\
log(\mu_i) & = \beta_1 \cdot W_{\text{T}[i]} + \beta_2 \cdot W_{\text{F}[i]}\\
& + \beta_3 \cdot PL_{W_\text{T}[i]} + \beta_4 \cdot PL_{W_\text{F}[i]}\\
\end{align}
$$

::: box-note
We allow the model to estimate the `PhonLev` effect for real words and nonce words separately:

-   $\beta_3$ is the effect of `PhonLev` when the word is real ($PL_{W_\text{T}}$).

-   $\beta_4$ is the effect of `PhonLev` in nonce words ($PL_{W_\text{F}}$).
:::

## Regression with interaction

```{r}
#| label: brm-int

brm_int <- brm(
  RT ~ 0 + IsWord + IsWord:PhonLev,
  data = mald,
  family = lognormal,
  cores = 4,
  seed = 9812,
  file = "data/cache/brm_int.rds"
)

brm_int
```

## RQ 3: Does the effect of mean phone-level distance differ depending on the lexical status?

```{r}
#| label: brm-int-draws

brm_int_draws <- as_draws_df(brm_int) |> 
  mutate(
    # Back ticks `` are necessary because the column names have colons ":"
    
    phonlev_diff = `b_IsWordFALSE:PhonLev` - `b_IsWordTRUE:PhonLev`
  )
```

## Difference in effect of `PhonLev` in nonce vs real words

```{r}
#| label: fig-phonlev-diff
#| fig-cap: "Predicted difference in `PhonLev` effect in nonce vs real words."
#| output-location: slide

brm_int_draws |> 
  ggplot(aes(phonlev_diff)) +
  stat_halfeye() +
  labs(
    x = "Difference (log)",
    y = element_blank()
  )
```

## CrIs of the difference in effect of `PhonLev`

```{r}
#| label: brm-int-cris
library(posterior)

brm_int_draws |> 
  summarise(
    lo_95 = quantile2(phonlev_diff, 0.025), hi_95 = quantile2(phonlev_diff, 0.975),
    lo_90 = quantile2(phonlev_diff, 0.5), hi_90 = quantile2(phonlev_diff, 0.95),
    lo_80 = quantile2(phonlev_diff, 0.1), hi_80 = quantile2(phonlev_diff, 0.9),
    lo_70 = quantile2(phonlev_diff, 0.15), hi_70 = quantile2(phonlev_diff, 0.85)
  ) |> 

  # Some pivot magic to make it prettier
  pivot_longer(everything()) |> 
  separate(name, c("boundary", "interval")) |> 
  pivot_wider(names_from = boundary) |> 
  mutate(lo = round(lo, 4), hi = round(hi, 4))
```

## Plot the model predictions

```{r}
#| label: brm-int-cond
conditional_effects(brm_int, "PhonLev:IsWord")
```

## Quantify the lexical status difference (log)

```{r}
#| label: phonlev-diff-5-10
#| warning: false

phonlev_diff_5_10 <- brm_int_draws |> 
  mutate(
    true_5 = b_IsWordTRUE + `b_IsWordTRUE:PhonLev` * 5,
    false_5 = b_IsWordFALSE + `b_IsWordFALSE:PhonLev` * 5,
    true_10 = b_IsWordTRUE + `b_IsWordTRUE:PhonLev` * 10,
    false_10 = b_IsWordFALSE + `b_IsWordFALSE:PhonLev` * 10,
    ft_5 = false_5 - true_5,
    ft_10 = false_10 - true_10
  ) |> 
  select(ft_5, ft_10)

phonlev_diff_5_10
```

## Plot the lexical status difference (log)

```{r}
#| label: fig-phonlev-diff-5-10

phonlev_diff_5_10 |> 
  pivot_longer(everything()) |> 
  ggplot(aes(value, name)) +
  stat_halfeye()
```

## Quantify the lexical status difference (ms)

```{r}
#| label: phonlev-diff-5-10-ms
#| warning: false

phonlev_diff_5_10_ms <- brm_int_draws |> 
  mutate(
    true_5 = exp(b_IsWordTRUE + `b_IsWordTRUE:PhonLev` * 5),
    false_5 = exp(b_IsWordFALSE + `b_IsWordFALSE:PhonLev` * 5),
    true_10 = exp(b_IsWordTRUE + `b_IsWordTRUE:PhonLev` * 10),
    false_10 = exp(b_IsWordFALSE + `b_IsWordFALSE:PhonLev` * 10),
    ft_5 = false_5 - true_5,
    ft_10 = false_10 - true_10
  ) |> 
  select(ft_5, ft_10)

phonlev_diff_5_10_ms
```

## Plot the lexical status difference (ms)

```{r}
#| label: fig-phonlev-diff-5-10-ms

phonlev_diff_5_10_ms |> 
  pivot_longer(everything()) |> 
  ggplot(aes(value, name)) +
  stat_halfeye()
```
